### 2019-07-04 2-Chapters/2-Chapter/chapter25.tex
First changes from Emilie's LONG comments
Cleanup empty pages between chapters etc. Less than 300 pages YAY

% A last solution is online algorithm selection, inspired from expert aggregation.
% Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292

\newpage  % WARNING ?

% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:25:conclusion}

In this chapter, we presented the first contribution of this manuscript,
that corresponds to our article \cite{Besson2018WCNC}.
%
We presented the use of aggregation algorithms,
especially for the real-world setting of unknown problem instances,
when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
% - \ExpQ and \Aggr works fine
% Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
Our proposed \Aggr{} was presented in details,
and we also highlighted its differences with \ExpQ.

% as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
%
% - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
We presented experiments on simple MAB problems, coming from previous work on bandits for OSA \cite{Jouini10},
and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
% It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
% and appears to be more robust than simple non-aggregated algorithms.

\ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
Empirically, \Aggr{} showed to always have a logarithmic
regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.

What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
%
%
Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
But this direction does not appear very useful for our objective at hand, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
% We will continue investigating regret bounds for \Aggr,
% and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).

% The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

\paragraph{Reproducility of the experiments.}
%
Experiments presented in this Section are based on our library SMPyBandits,
and the page \href{https://SMPyBandits.GitHub.io/Aggregation.html}{\texttt{SMPyBandits.GitHub.io/Aggregation.html}} gives instructions to reproduce them.

>>>>

% % A last solution is online algorithm selection, inspired from expert aggregation.
% % Include here the discussion about expert aggregation and my \textbf{Aggregator} algorithm, see https://hal.inria.fr/hal-01705292

% \newpage  % WARNING ?

% % ----------------------------------------------------------------------------
% \section{Conclusion}
% \label{sec:25:conclusion}

% In this chapter, we presented the first contribution of this manuscript,
% that corresponds to our article \cite{Besson2018WCNC}.
% %
% We presented the use of aggregation algorithms,
% especially for the real-world setting of unknown problem instances,
% when tuning parameters before-hand is no longer possible and an adaptive algorithm is preferable.
% % - \ExpQ and \Aggr works fine
% % Both algorithms \ExpQ{} and \Aggr{} have been detailed, and we explained why
% Our proposed \Aggr{} was presented in details,
% and we also highlighted its differences with \ExpQ.

% % as well \color{red} as the intuition that it seems more suited for purely stochastic problems\color{black}.
% %
% % - \Aggr{} really help to select the best algorithm against a certain problem, on the fly without any prior knowledge of the problem neither any prior knowledge on which algorithms is the best
% We presented experiments on simple MAB problems, coming from previous work on bandits for OSA \cite{Jouini10},
% and the simulations results showed that \Aggr{} is able to identify on the fly the best algorithm to trust for a specific problem, as expected.
% Experiments on problems mixing different families of distributions were also presented, with similar conclusions in favor of \Aggr.
% % It is not presented in this section, but our proposed algorithm also works well in dynamic scenarios, in which the distribution of the arms can change abruptly at some time,
% % and appears to be more robust than simple non-aggregated algorithms.

% \ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
% Empirically, \Aggr{} showed to always have a logarithmic
% regret if it aggregates at least one algorithm with logarithmic regrets (like UCB, \klUCB, Thompson sampling, BayesUCB etc).
% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
% As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this is an over-optimistic goal.

% What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
% %
% %
% Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
% But this direction does not appear very useful for our objective at hand, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
% % We will continue investigating regret bounds for \Aggr,
% % and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).

% % The numerical simulations used our Python open-source library, SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% % The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
Cleanup empty pages between chapters etc. Less than 300 pages YAY

We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from $2019$) and are usually more complex, or based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

\paragraph{Future works.}
%
An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
and even if some have already been implemented, a major future work is to implement the most interesting ones
(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).

Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits of \cite{Seznec2018} (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

Finally, the most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used SMPyBandits.
While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
%
We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to our work.

As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
I would also love to be able to teach a graduate or under-graduate course on multi-armed bandits in the next years, and use my library as a support to illustrate the course, as well as practical sessions.

@@

We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
We conclude by a discussion on parallel multi-core CPU computations to speed-up SMPyBandits.

\subsection{Minimalist example of use of SMPyBandits}
\label{sub:3:miniExampleSMPyBandits}

Below is presented in Code~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
Only the mean reward is printed at the end of the simulation.
%
With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
\begin{verbatim}
The UCB algorithm obtains here a mean reward = 0.886
\end{verbatim}
% \vspace*{-10pt} % WARNING
%
% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
    % \label{lst:3:exampleOfUse}
% \end{listing}
\end{small}

\subsection{A minimalist \texttt{main.py} file}
\label{sub:3:miniMainFileSMPyBandits}

We present below in Code~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
    % \caption{Small example of \texttt{main.py} file}
    \captionof{listing}{Example of \texttt{main.py} file\label{lst:3:main}.}
    % \label{lst:3:main}
% \end{listing}
\end{small}

\subsection{A minimalist \texttt{configuration.py} file}
\label{sub:3:miniConfigurationFileSMPyBandits}

We present below in Code~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
which is used to define the experiment.
It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:

% \begin{verbatim}
%     (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
% \end{verbatim}

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
    \end{minted}
    \caption{Example of Bash code to run an experiment.}
    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
    \label{lst:3:howToRunExperiment2}
\end{listing}

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
    % \caption{Small example of \texttt{configuration.py} file}
    \captionof{listing}{Example of \texttt{configuration.py} file\label{lst:3:configuration}.}
    % \label{lst:3:configuration}
% \end{listing}
\end{small}

@@

\begin{enumerate}
    \setcounter{enumi}{9}
    \item
    % Algorithm 10 is
    $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}).
    It is a recent variant of the $\mathrm{MOSS}$ algorithm \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter. It is anytime, and obtains good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
    Empirically, it performs usually similarly to \UCB, and is slightly more costly in terms of both time and memory.

    \item
    % Algorithm 11 is
    Approximated Finite-Horizon Gittins from \cite{Lattimore16a} (\texttt{ApproximatedFHGittins}), using $\alpha=1$.
    It mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in terms of both time and memory.

    \item
    % Algorithm 12 is
    $\UCB\dagger$ from the same author \cite{Lattimore2018refining} (\texttt{UCBdagger}), using $\alpha=1$.
    It mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
    It also achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in computation time, but it is much more costly in terms of memory.

    \item
    % Algorithm 13 is
    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}
    (\texttt{klUCBswitchAnytime}),
    is a very recent variant of the \KLUCB{} algorithm \cite{KLUCBJournal}.
    It mimics the \KLUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
    It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
    Empirically, it usually outperforms (slightly) \klUCB{} in terms of regret, and costs about the same memory, but it is slower.

@@

    \item
    % Algorithm 15 is
    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019} (\texttt{RCB}), using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$. We prefer to use the simplest of the algorithms described in the article, and other algorithms in the same article used other families of distributions for the perturbations.
    It was analyzed and found to be order-optimal for problem-dependent bounds, and empirically we found that it is usually performs slightly worse than \UCB, in terms of regret, time and memory.
    It is included because it remains comparably efficient with the other state-of-the-art algorithms, and because the key message from \cite{KimTewari2019} is very interesting:
    ``the optimism embedded in UCB can be replaced by simple randomization''.

@@

\paragraph{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.

@@

\paragraph{Comparing measurements.}
But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
Relative difference for costs do not depend much on the code and the machine used for the simulation.
%
% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
%
Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
%
We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.

\paragraph{Normalizing data.}
%
The two figures below regret normalized data, in the following way.
For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
%
We know that efficient algorithms are expected to follow these patterns:
\begin{itemize}%\tightlist
    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
\end{itemize}
Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
%
On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).

>>>>

We focus on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from $2019$) and are usually more complex, or based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

\paragraph{Future works.}
%
% An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
% For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
% For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
% or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
% %
% At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
% and even if some have already been implemented, a major future work is to implement the most interesting ones
% (tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).

% Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
% Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits of \cite{Seznec2018} (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

The most exciting thing that could happen to our SMPyBandits library would be to see it gaining popularity!
Its documentation has seen already about $20000$ visits, the project on GitHub had $88$ stars in June $2019$, and based on the features requested and the emails received about it, we counted between $10$ to $20$ researchers in other labs who use or used SMPyBandits.
While this is a good start, we believe that the library is mature and interesting enough to hope to see it being used by more people.
%
We have submitted a summary paper presenting SMPyBandits \cite{SMPyBanditsJMLR} to the MLOSS track of the Journal of Machine Learning Research (JMLR), and we hope that it will be accepted, as it would give more visibility to our work.
%
As a personal note, I would like to continue working on SMPyBandits, and implement the most important requested features, as well as maintain it, and possibly continue to add recent algorithms by following actively the research in this community.
I would also love to be able to teach a graduate or under-graduate course on multi-armed bandits in the next years, and use my library as a support to illustrate the course, as well as practical sessions.

@@

% We start by giving a small but complete example of use of SMPyBandits, then we give additional experimental results completing those presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above.
We give in this appendix more details regarding the numerical experiments presented in Section~\ref{sec:3:reviewSPAlgorithms} and \ref{sec:3:timeAndMemoryCosts} above,
for the other algorithms used in our benchmark,
and details on the methodology used to measure the time and memory.
% We conclude by a discussion on parallel multi-core CPU computations to speed-up SMPyBandits.

% \subsection{Minimalist example of use of SMPyBandits}
% \label{sub:3:miniExampleSMPyBandits}

% Below is presented in Code~\ref{lst:3:exampleOfUse} a minimalist example of use of the library, to define a toy bandit problem with $K=2$ Bernoulli distributed arms (of means $\mu=0.1$ and $\mu=0.9$), one \UCB{} algorithm, and play one experiment of horizon $T=1000$.
% Only the mean reward is printed at the end of the simulation.
% %
% With Python 3.6 this example shows that the algorithm got a mean reward very close to the optimal arm, $\hat{\mu} = 0.886 \simeq 0.9 = \mu^* = \mu_2$.
% \begin{verbatim}
% The UCB algorithm obtains here a mean reward = 0.886
% \end{verbatim}
% % \vspace*{-10pt} % WARNING
% %
% % https://tex.stackexchange.com/a/12430/
% \begin{small}
% % \begin{listing}[h!]
%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_use_of_SMPyBandits.py}
%     % \caption{Small example of \texttt{main.py} file}
%     \captionof{listing}{Example of use of SMPyBandits\label{lst:3:exampleOfUse}.}
%     % \label{lst:3:exampleOfUse}
% % \end{listing}
% \end{small}

% \subsection{A minimalist \texttt{main.py} file}
% \label{sub:3:miniMainFileSMPyBandits}

% We present below in Code~\ref{lst:3:main} a full example of a minimalist \texttt{main.py} file,
% which is used to load the configuration, run the experiment and then print and plot various visualizations for the simulation.

% % https://tex.stackexchange.com/a/12430/
% \begin{small}
% % \begin{listing}[h!]
%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_main_singleplayer.py}
%     % \caption{Small example of \texttt{main.py} file}
%     \captionof{listing}{Example of \texttt{main.py} file\label{lst:3:main}.}
%     % \label{lst:3:main}
% % \end{listing}
% \end{small}

% \subsection{A minimalist \texttt{configuration.py} file}
% \label{sub:3:miniConfigurationFileSMPyBandits}

% We present below in Code~\ref{lst:3:configuration} a full example of a minimalist \texttt{configuration.py} file,
% which is used to define the experiment.
% It defines the number of arms $K$ and their distribution, the horizon $T$, the policies to test etc.
% It uses environment variables, so a user can configure some parameters of the simulation when launching them from a terminal:

% % \begin{verbatim}
% %     (bash) $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
% % \end{verbatim}

% % https://tex.stackexchange.com/a/12430/
% \begin{listing}[h!]
%     \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
% $ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
%     \end{minted}
%     \caption{Example of Bash code to run an experiment.}
%     % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
%     \label{lst:3:howToRunExperiment2}
% \end{listing}

% % https://tex.stackexchange.com/a/12430/
% \begin{small}
% % \begin{listing}[h!]
%     \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_configuration_singleplayer.py}
%     % \caption{Small example of \texttt{configuration.py} file}
%     \captionof{listing}{Example of \texttt{configuration.py} file\label{lst:3:configuration}.}
%     % \label{lst:3:configuration}
% % \end{listing}
% \end{small}

@@

\begin{enumerate}
    \setcounter{enumi}{9}
    \item
    % Algorithm 10 is
    $\mathrm{MOSS}$-$\mathrm{Anytime}$ from \cite{Degenne16}, using $\alpha=1$ (\texttt{MOSSAnytime}),
    is an anytime variant of $\mathrm{MOSS}$ \cite{BubeckSlivkins12}, using an adaptive tuning of its parameter. It obtains good ``best of both world'' performances, meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
    Empirically, it performs usually similarly to \UCB, and is slightly more costly in terms of both time and memory.

    \item
    % Algorithm 11 is
    Approximated Finite-Horizon Gittins from \cite{Lattimore16a} (\texttt{ApproximatedFHGittins}), using $\alpha=1$,
    mimics the \UCB{} algorithm but uses a more complicated function to compute the upper confidence bounds.
    It achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in terms of both time and memory.

    \item
    % Algorithm 12 is
    $\UCB\dagger$ from the same author \cite{Lattimore2018refining} (\texttt{UCBdagger}), using $\alpha=1$,
    mimics the \UCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
    It also achieves order-optimal problem-dependent bounds with a $\cO(K\log(T)/\Delta^2)$ regret.
    Empirically, it performs usually similarly to \UCB{} and $\mathrm{MOSS}$-$\mathrm{Anytime}$, and is slightly more costly in computation time, but it is much more costly in terms of memory.

    \item
    % Algorithm 13 is
    Anytime variant of KL-UCB-Switch from \cite{GarivierHadiji2018}
    (\texttt{klUCBswitchAnytime}),
    is a very recent variant of \KLUCB{} \cite{KLUCBJournal}.
    It mimics the \KLUCB{} algorithm but uses a much more complicated function to compute the upper confidence bounds, and uses more storage.
    It is anytime, and was prove to obtain good ``best of both world'' performances, meaning that it achieves an asymptotically optimal $\cO(\log(T))$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.
    Empirically, it usually outperforms (slightly) \klUCB{} in terms of regret, and costs about the same memory, but it is slower.

@@

    \item
    % Algorithm 15 is
    $\mathrm{RCB}$ (Randomized Confidence Bound) from \cite{KimTewari2019} (\texttt{RCB}), using $\alpha=1$ and perturbations uniformly sampled in $[0,1]$. We prefer to use the simplest of the algorithms described in the article.
    %  and other algorithms in the same article used other families of distributions for the perturbations.
    It was analyzed and found to be order-optimal for problem-dependent bounds, and empirically we found that it is usually performs slightly worse than \UCB, in terms of regret, time and memory.
    It is included because it remains comparably efficient with the other state-of-the-art algorithms, and because the key message from \cite{KimTewari2019} is very interesting:
    ``the optimism embedded in UCB can be replaced by simple randomization''.

@@

\textbf{About the measurement protocol.}
%
To be precise, we used the two following approaches to measure the real time and memory costs of the different algorithms.

@@

\textbf{Comparing measurements.}
%
But what is important in these simulation results is not the exact values, but rather to compare the costs of the different algorithms, and thus only the relative difference in terms of time and memory costs are important.
Relative difference for costs do not depend much on the code and the machine used for the simulation.
%
% The tables~\ref{table:3:time_problem1}, \ref{table:3:time_problem2} \ref{table:3:memory_problem1} and \ref{table:3:memory_problem2} below follow the same structure as the ones given in Section~\ref{sec:3:reviewSPAlgorithms} but report results in terms of computation time (in milli-seconds) and memory (in B).
The computation time is \emph{normalized}, that is it is divided by the horizon, to count the (average) time used by an algorithm at time $t$ from $t=1$ to $t=T$, while the memory cost is \emph{not normalized}.
%
Thus, we can check that efficient algorithms have a running time independent on the horizon $T$, instead of just observing a linear dependency, and we can check that efficient algorithms have a memory cost independent on the horizon.
All the considered algorithms in this Chapter are efficient in this aspect, but in Chapter~\ref{chapter:6} we study algorithms that essentially have to store, and run computations, on an increasing number of observations (\eg, \CUSUMUCB), so their computation cost at time $t$ increases (linearly) with $t$, and thus they are more costly.
%
We can also check that efficient algorithms have a running time and a memory cost linear in the number of arms $K$, while more complex algorithms like BESA suffer from an exponential blow-up of their complexity when $K$ increases.

\textbf{Normalizing data.}
%
The two figures below regret normalized data, in the following way.
For each algorithm, we ran simulations to obtain its (empirical) regret, computation time and space requirement, for different problems with different values of $K$ and $T$.
Simply considering the average of such measurements makes no sense, as for instance two values of the regret for $T=1000$ or $T=50000$ do not have the same order of magnitude.
%
We know that efficient algorithms are expected to follow these patterns:
\begin{itemize}%\tightlist
    \item The final regret $R_T$ should scale as $\cO(K \log(T))$,
    \item The total computation time $\cT_T$ should scale as $\cO(K T)$ (\ie, a computation time of $\cO(K)$ for each time step $t$),
    \item The total memory cost $\cM_T$ should scale as $\cO(K)$, independent of the horizon.
\end{itemize}
Thus, on the one hand, when we show aggregated results from all the different values of $K$ and $T$, we normalized the data by dividing the regrets by $K \log(T)$, the times by $K T$ and the memory by $K$.
%
On the other hand, when we plot a quantity ($R_T$, $\cT_T$, $\cM_T$) as a function of $T$ (resp. of $K$) we only normalize by $K$ (resp. by $T$).

### 2019-07-04 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
First changes from Emilie's LONG comments
Cleanup empty pages between chapters etc. Less than 300 pages YAY

% \paragraph{Note on the simulation code.}
\paragraph{Reproducility of the experiments.}
%
The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License, and published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in July and August $2018$.

>>>>

% \paragraph{Note on the simulation code.}
\paragraph{Reproducibility.}
%
The source code (MATLAB or Octave) used for the simulations and the figures of this section is open-sourced under the MIT License.
It was written in collaboration with Rémi Bonnefoi and Julio César Mango-Vasquez, in the summer $2018$,
and is published at \href{https://Bitbucket.org/scee_ietr/ucb_smart_retrans}{\texttt{Bitbucket.org/scee\_ietr/ucb\_smart\_retrans}}.
